---
title: "Neighborhood physical characteristics and poor self-rated health among older Brazilian adults: The ELSI-Urbe Project"
author: "Amanda Silva Magalhães; Amanda Cristina de Souza Andrade; Bruno de Souza Moreira; Maria Fernanda Lima-Costa; Alex Quistberg; Waleska Teixeira Caiaffa"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-location: left
    number-sections: true
    code-fold: true
    page-layout: full
execute:
  freeze: auto
  cache: true
---



```{r}
#| label: Packages
#| code-summary: Packages
#| warning: false
#| message: false

rm(list = ls())

library(pacman)
p_load(
  # Data manipulation and wrangling
  tidyverse, dplyr, purrr, here,
  # Data import
  haven,
  # Survey data analysis
  survey, srvyr,
  # Statistical modeling
  glmmTMB, car, sandwich, lmtest, lme4,
  # Model outputs and tidying
  broom.mixed, broom.helpers,
  # Tables and reporting
  gt, gtsummary, kableExtra,
  # Causal inference and DAGs
  dagitty, ggdag
)
```

```{r}
#| label: Database
#| code-summary: Database

df <- read_dta(here("elsi_brasil_auditoria_virtual.dta"))
```



# Introduction

Self-rated health is a key predictor of morbidity, mortality, and health care utilization. Neighborhood physical characteristics are associated with self-rated health, particularly among older adults, who tend to reside in the same area longer and rely on local resources. This study investigated the association between neighborhood physical characteristics, assessed through virtual audits, and poor self-rated health among older adults in Brazil.

# Methods

## ELSI-Urbe

The present research is part of the project “The Influence of the Physical and Social Environment on the Health of Older Brazilian Adults: A Longitudinal Multi-Method Population-Based Study” (ELSI-Urbe), initiated in 2021 by the Observatory for Urban Health in Belo Horizonte (OSUBH) at the Federal University of Minas Gerais (UFMG), in partnership with the Center for Studies in Public Health and Aging at UFMG and the René Rachou Institute – Fiocruz Minas of the Oswaldo Cruz Foundation.

The ELSI-Urbe project investigates how physical and social neighborhood characteristics influence the health and well-being of older adults over time. It is part of the broader Brazilian Longitudinal Study of Aging (ELSI-Brazil) and uses data from its participants to understand the long-term effects of urban environments on aging ([Coelho et al., 2025](https://doi.org/10.1007/s11524-025-00998-5)). The project integrates individual-level data from ELSI-Brazil with ultra-local contextual information from secondary sources and virtual audits, aggregated across different geographic units. Only participants living in urban areas were included.

## ELSI-Brazil

The ELSI-Brazil is a nationally representative, population-based cohort study of community-dwelling adults aged 50 years or older, living in 70 municipalities across the five Brazilian regions. The sampling process was conducted in stages, combining the stratification of primary sampling units (municipalities), census tracts, and households. The first wave was conducted in 2015–2016, the second in 2019–2021, and the third in 2023–2024. All residents aged 50 years or older in selected households were eligible to participate in interviews using a structured questionnaire and physical measurements. Further details about ELSI-Brazil are available in previous publications ([Lima-Costa et al., 2018](https://doi.org/10.1093/aje/kwx387); [Lima-Costa et al., 2023](https://doi.org/10.1093/ije/dyac132)) and on the [ELSI-Brazil website](https://elsi.cpqrr.fiocruz.br/en/home-english/).

## Virtual audit

The ELSI-Urbe Virtual Audit subproject was conducted using a systematic observation method between April and October 2024, based on available Google Street View images. The audit was performed at the street-segment level, considering both sides of the street, and later aggregated into census tracts, which were defined as neighborhoods. The sampling process was carried out in stages, sequentially selecting municipalities (the most populous municipality participating in ELSI-Brazil from each federative unit, with 20 available), census tracts (with ELSI-Brazil interviews), and random street segments. Data collection was conducted by a single trained rater using a multidimensional instrument and a field manual.

## Study sample

The analytical sample followed a two-level multilevel structure in which individuals from the second wave of ELSI-Brazil (level 1) were nested within residential census tracts (level 2). The eligible population comprised participants residing in urban areas whose census tracts had available virtual audit data (n = 2,794 individuals nested within 218 census tracts).

## Outcome

The outcome variable was self-rated health, measured with the question: “In general, how do you rate your health?” Response options were categorized as 1 = poor (regular/poor/very poor) and 0 = good (excellent/very good/good). The final analytical sample considering the outcome was 2,766 individuals.

## Exposures

The exposure variables consisted of six indicators derived from the virtual audit data. Composite indicators were created based on urban infrastructure, traffic, and physical disorder, resulting in two components for each of the three domains assessed, as defined in a previous study.

For the urban infrastructure domain, the first subdomain, pathway accessibility, included the items “street condition,” “sidewalk extends along the entire segment,” and “unobstructed sidewalk.” The second subdomain, surface and continuity, consisted of the items “street material,” “presence of sidewalks,” and “sidewalks even along the entire segment” (total variance: 59.7%; Cronbach’s alpha: 0.672).

In the traffic domain, the first subdomain, markings and controls, comprised the items “lane markings,” “median lane,” “traffic control devices,” “crosswalk,” and “public transport stop.” The second subdomain, speed and bus lane, included the items “speed control devices” and “dedicated bus lane” (total variance: 48.8%; Cronbach’s alpha: 0.675).

In the physical disorder domain, the first subdomain, absence of infrastructure and vegetation, was represented by the items “condition of electrical wires,” “absence of trees,” and “absence of gardens or potted plants.” The second subdomain, absence of buildings and green spaces, consisted of the items “buildings in poor condition” and “absence of parks or green spaces” (total variance: 61.1%; Cronbach’s alpha: 0.617).

## Adjustment variables

Adjustment variables initially included sociodemographic characteristics, namely sex (female and male), age group (50–59, 60–69, 70–79, and ≥80 years), race/color (white, black or brown, and other (asian and indigenous)), education (≤3, 4–7, and ≥8 years of schooling), marital status (unmarried and married or cohabiting), number of household residents (1, 2, 3, and ≥4), and household income per capita in tertiles (≤933; >933–1,500; and >1,500 reais), as well as health and lifestyle characteristics, including number of chronic diseases (0, 1, and ≥2), depressive symptoms (no and yes), fruit and vegetable consumption (<5 and ≥5 days/week), global physical activity (inactive and active), binge drinking (no and yes), and current tobacco use (no and yes).

However, based on the directed acyclic graph (DAG) and the minimal sufficient adjustment set identified, the final models were adjusted for sex, age group, race/color, education, and household income per capita only.

```{r, fig.width=16, fig.height=10, out.width="100%"}
#| label: Directed acyclic graph (DAG)
#| code-summary: Directed acyclic graph (DAG)

## Variable definitions
# Exposure: physical_environment
# Outcome: poor_self_rated_health

## Confounders
# These variables are considered common causes of both
# physical_environment and poor_self_rated_health:
# sex, age_group, race_color, education, and household_income_per_capita.

## Additional health-related covariates
# These variables are associated with poor_self_rated_health:
# marital_status, number_household_residents, number_chronic_diseases,
# fruit_vegetable_consumption, binge_drinking, and current_tobacco_use.

## Potential mediators
# These variables are hypothesized to lie on the causal pathway between
# physical_environment and poor_self_rated_health:
# global_physical_activity and depressive_symptoms.

dag <- dagitty("dag {

physical_environment [exposure]
poor_self_rated_health [outcome]

physical_environment -> poor_self_rated_health

sex -> physical_environment
sex -> poor_self_rated_health

age_group -> physical_environment
age_group -> poor_self_rated_health

race_color -> physical_environment
race_color -> poor_self_rated_health

education -> physical_environment
education -> poor_self_rated_health

household_income_per_capita -> physical_environment
household_income_per_capita -> poor_self_rated_health

marital_status -> poor_self_rated_health
number_household_residents -> poor_self_rated_health
number_chronic_diseases -> poor_self_rated_health
fruit_vegetable_consumption -> poor_self_rated_health
binge_drinking -> poor_self_rated_health
current_tobacco_use -> poor_self_rated_health

physical_environment -> global_physical_activity
global_physical_activity -> poor_self_rated_health

physical_environment -> depressive_symptoms
depressive_symptoms -> poor_self_rated_health

}")

ggdag(dag, layout = "stress") +
  geom_dag_edges(edge_colour = "grey35") +
  geom_dag_node(color = "grey95", size = 16, shape = 16) + 
  geom_dag_label(aes(label = name), 
                 fill = "white",
                 color = "black",
                 label.size = 0,
                 size = 4,
                 alpha = 0.8) +
  theme_dag() +
  theme(panel.background = element_rect(fill = "white"),
        plot.title = element_blank())
```

```{r}
#| label: Minimum sufficient adjustment set
#| code-summary: Minimum sufficient adjustment set

adjustmentSets(dag)
```

## Statistical analysis

Descriptive statistics were calculated using frequency distributions (%), means, and standard deviations (SD), accounting for the complex sampling design and individual weights using the survey package. Associations between exposures and outcomes were examined using prevalence ratios (PR) and 95% confidence intervals (CI), estimated using Poisson regression models with a log link and robust variance clustered at the census tract level. The models were adjusted for sociodemographic characteristics. Normalized individual sampling weights were incorporated into the regression models. Multicollinearity among independent variables was assessed using the variance inflation factor (VIF). Post-estimation diagnostics were performed, including assessment of dispersion, Pearson residuals, Cook’s distance, and leverage to evaluate model fit and identify potential outliers and influential observations. The significance level was set at 5%. All analyses were conducted using R software (R version 4.5.0; R Development Core Team).



# Results


## Descriptive analyses

```{r}
#| label: Distribution of individuals by area of residence
#| code-summary: Distribution of individuals by area of residence

df %>%
  select(zona) %>%
  mutate(
    zona = factor(zona, levels = c(0, 1), labels = c("Rural", "Urban"))
  ) %>%
  tbl_summary(
    label = list(zona ~ "Area"),
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_categorical() ~ c(0, 1)),
    missing_text = "Unknown"
  ) %>%
  modify_header(
    label = "**Characteristic**",
    stat_0 = "**Total**<br>n = {N}"
  ) %>%
  modify_footnote(everything() ~ NA)
```

```{r}
#| label: Distribution of cities and virtual urban audit status
#| code-summary: Distribution of cities and virtual urban audit status

df %>%
  filter(zona == 1) %>%
  distinct(cidade2, .keep_all = TRUE) %>% 
  mutate(
    cidade_auditoria = factor(cidade_auditoria, 
                              levels = c(0, 1), 
                              labels = c("No", "Yes"))
  ) %>%
  select(cidade_auditoria) %>%
  tbl_summary(
    label = list(cidade_auditoria ~ "Virtual audit"),
    type = list(cidade_auditoria ~ "categorical"),
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(
    label = "**Characteristic**",
    stat_0 = "**Total cities**<br>n = {N}"
  ) %>%
  modify_footnote(everything() ~ NA)
```

```{r}
#| label: Cities and virtual urban audit status
#| code-summary: Cities and virtual urban audit status

df %>%
  filter(zona == 1) %>%
  distinct(cidade2, cidade_auditoria) %>%
  arrange(cidade_auditoria, cidade2) %>%
  mutate(
    `#` = row_number(),
    cidade_auditoria = if_else(cidade_auditoria == 1, "Yes", "No")
  ) %>%
  select(`#`, everything()) %>%
  knitr::kable(
    col.names = c("#", "City", "Virtual audit"))
```

```{r}
#| label: Distribution of census tracts and virtual audit status
#| code-summary: Distribution of census tracts and virtual audit status

df %>%
  filter(zona == 1) %>%
  distinct(setor, .keep_all = TRUE) %>% 
  mutate(
    setor_auditoria = factor(
      setor_auditoria,
      levels = c(0, 1),
      labels = c("No", "Yes"))
  ) %>%
  select(setor_auditoria) %>%
  tbl_summary(
    label = list(setor_auditoria ~ "Virtual audit"),
    type = list(setor_auditoria ~ "categorical"),
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(
    label = "**Characteristic**",
    stat_0 = "**Total census tracts**<br>n = {N}"
  ) %>%
  modify_footnote(everything() ~ NA)
```

```{r}
#| label: Distribution of individuals in census tracts and virtual urban audit status
#| code-summary: Distribution of individuals in census tracts and virtual urban audit status

df %>%
  filter(zona == 1) %>%
  mutate(
    setor_auditoria = factor(
      setor_auditoria,
      levels = c(0, 1),
      labels = c("No", "Yes"))
  ) %>%
  select(setor_auditoria) %>%
  tbl_summary(
    label = list(setor_auditoria ~ "Virtual audit"),
    type = list(setor_auditoria ~ "categorical"),
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(
    label = "**Characteristic**",
    stat_0 = "**Total**<br>n = {N}"
  ) %>%
  modify_footnote(everything() ~ NA)
```


## Self-rated health

```{r}
#| label: Excellent/Very good/Good vs Regular/Poor/Very poor
#| code-summary: Excellent/Very good/Good vs Regular/Poor/Very poor

options(survey.lonely.psu = "adjust")

df %>%
  filter(zona == 1, !is.na(identificador), !is.na(aas_inv2)) %>%
  mutate(
    aas_inv2 = factor(
      aas_inv2,
      levels = c(0, 1),
      labels = c("Good", "Poor"))
  ) %>%
  as_survey_design(
    ids = upa,
    strata = estrato,
    weights = peso_calibrado,
    nest = TRUE
  ) %>%
  select(aas_inv2) %>%
  tbl_svysummary(
    label = list(aas_inv2 ~ "Self-rated health"),
    statistic = list(all_categorical() ~ "{p}%"),
    digits = list(all_categorical() ~ 1)
  ) %>%
  modify_header(
    label = "**Characteristic**",
    stat_0 = "**Total**"
  ) %>%
  modify_footnote(everything() ~ NA)
```


## Individual and neighborhood characteristics stratified by self-rated health

```{r}
#| label: Individual and neighborhood characteristics stratified by self-rated health
#| code-summary: Individual and neighborhood characteristics stratified by self-rated health
#| warning: false
#| message: false
#| cache: true
#| cache.lazy: false

options(survey.lonely.psu = "adjust")

analysis_df <-
  df %>%
  filter(zona == 1, !is.na(identificador), !is.na(aas_inv2)) %>%
  mutate(
    aas_inv2 = haven::as_factor(aas_inv2),
    sexo = factor(sexo, levels = c(0, 1), labels = c("Female", "Male")),
    faixaetaria = factor(
      faixaetaria,
      levels = c(1, 2, 3, 4),
      labels = c("50-59", "60-69", "70-79", "≥80")),
    cor_raca2 = factor(
      cor_raca2,
      levels = c(1, 2, 3),
      labels = c("White", "Black or brown", "Other")),
    escolaridade2 = factor(
      escolaridade2,
      levels = c(1, 2, 3),
      labels = c("≤3", "4-7", "≥8")),
    estado_civil = factor(
      estado_civil,
      levels = c(0, 1),
      labels = c("Unmarried", "Married or cohabiting")),
    nmoradores2 = factor(
      nmoradores2,
      levels = c(1, 2, 3, 4),
      labels = c("1", "2", "3", "≥4")),
    rendadompc2corr = factor(
      rendadompc2corr,
      levels = c(1, 2, 3),
      labels = c("1st tertile: ≤933", "2nd tertile: >933-1,500", "3rd tertile: >1,500")),
    dcnt2 = factor(
      dcnt2,
      levels = c(0, 1, 2),
      labels = c("0", "1", "≥2")),
    sint_dep = factor(
      sint_dep,
      levels = c(0, 1),
      labels = c("No", "Yes")),
    regularFH = factor(
      regularFH,
      levels = c(0, 1),
      labels = c("<5", "≥5")),
    af_global = factor(
      af_global,
      levels = c(0, 1),
      labels = c("Inactive", "Active")),
    consumo_abusivo = factor(
      consumo_abusivo,
      levels = c(0, 1),
      labels = c("No", "Yes")),
    consumo_tabaco = factor(
      consumo_tabaco,
      levels = c(0, 1),
      labels = c("No", "Yes")))

n_total <- nrow(analysis_df)
n_good <- sum(analysis_df$aas_inv2 == "Excelente/Muito boa/Boa", na.rm = TRUE)
n_poor <- sum(analysis_df$aas_inv2 == "Regular/Ruim/Muito ruim", na.rm = TRUE)

analysis_df %>%
  as_survey_design(
    ids = upa,
    strata = estrato,
    weights = peso_calibrado,
    nest = TRUE
  ) %>%
  tbl_svysummary(
    by = aas_inv2,
    include = c(
      sexo, faixaetaria, cor_raca2, escolaridade2, estado_civil, nmoradores2, rendadompc2corr,
      dcnt2, sint_dep, regularFH, af_global, consumo_abusivo, consumo_tabaco,
      dom_urbana1_escala, dom_urbana2_escala,
      dom_transito1_escala, dom_transito2_escala,
      dom_desordem1_escala, dom_desordem2_escala),
    type = list(
      sint_dep ~ "categorical",
      consumo_abusivo ~ "categorical",
      consumo_tabaco ~ "categorical"),
    label = list(
        sexo ~ "Sex",
        faixaetaria ~ "Age group (years)",
        cor_raca2 ~ "Race/Color",
        escolaridade2 ~ "Education (years)",
        estado_civil ~ "Marital status",
        nmoradores2 ~ "Number of household residents",
        rendadompc2corr ~ "Household income per capita in tertiles (R$)",
        dcnt2 ~ "Number of chronic diseases",
        sint_dep ~ "Depressive symptoms,",
        regularFH ~ "Fruit and vegetable consumption",
        af_global ~ "Global physical activity",
        consumo_abusivo ~ "Binge drinking",
        consumo_tabaco ~ "Current tobacco use",
        dom_urbana1_escala ~ "Pathway accessibility [mean (SD)]",
        dom_urbana2_escala ~ "Surface and continuity [mean (SD)]",
        dom_transito1_escala ~ "Markings and controls [mean (SD)]",
        dom_transito2_escala ~ "Speed and bus lane [mean (SD)]",
        dom_desordem1_escala ~ "Absence of infrastructure and vegetation [mean (SD)]",
        dom_desordem2_escala ~ "Absence of building and green spaces [mean (SD)]"),
    statistic = list(
      all_categorical() ~ "{p}",
      all_continuous() ~ "{mean} ({sd})"),
    digits = list(
      all_categorical() ~ 1,
      all_continuous() ~ 1),
    missing = "no"
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(
    pvalue_fun = ~ style_pvalue(.x, digits = 3)
  ) %>%
  add_ci(
    style_fun = list(
      all_categorical() ~ function(x) sprintf("%.1f", x * 100),
      all_continuous() ~ function(x) style_sigfig(x, digits = 1))
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(
        dplyr::across(
          starts_with("ci_stat"),
          ~ ifelse(
            variable %in% c(
              "dom_urbana1_escala",
              "dom_urbana2_escala",
              "dom_transito1_escala",
              "dom_transito2_escala",
              "dom_desordem1_escala",
              "dom_desordem2_escala"
            ),
            "",
            .x
          )
        )
      )
  ) %>%
  modify_fmt_fun(
    update = list(
      all_stat_cols() ~ function(x) gsub("%", "", x),
      starts_with("ci_stat") ~ function(x) {
        x <- gsub("%", "", x)
        x <- gsub(", ", " - ", x)
        ifelse(x == "" | is.na(x), "", x)
      }
    )
  ) %>%
  modify_header(
    label = "**Characteristic**",
    all_stat_cols() ~ "**%**",
    starts_with("ci_stat") ~ "**95% CI**",
    p.value = "**p-value**"
  ) %>%
  modify_spanning_header(
    c(stat_0, ci_stat_0) ~ paste0("**Total<br>n=", n_total, "**"),
    c(stat_1, ci_stat_1) ~ paste0("**Good self-rated health<br>n=", n_good, "**"),
    c(stat_2, ci_stat_2) ~ paste0("**Poor self-rated health<br>n=", n_poor, "**")
  ) %>%
  modify_footnote(everything() ~ NA) %>%
  as_gt() %>%
  gt::tab_source_note(
    source_note = "Note: Missing values for Race/Color (n=6); Education (n=36); Household income per capita (n=163); Depressive symptoms (n=321); Fruit and vegetable consumption (n=30); Global physical activity (n=62); Binge drinking (n=10). P-values were calculated using the Rao–Scott adjusted chi-square test for categorical variables and the survey-weighted Wald test for continuous variables."
  )
```


## Multilevel poisson models with robust variance clustered at census tract

```{r}
#| label: Crude association
#| code-summary: Crude association
#| warning: false
#| message: false

analysis_data_ct_crude <- df %>%
  filter(zona == 1, !is.na(identificador), !is.na(aas_inv2)) %>%
  mutate(
    census_tract = factor(setor),
    normalized_weight = peso_calibrado / mean(peso_calibrado, na.rm = TRUE))

exposure_labels <- c(
  dom_urbana1_escala = "Pathway accessibility",
  dom_urbana2_escala = "Surface and continuity",
  dom_transito1_escala = "Markings and controls",
  dom_transito2_escala = "Speed and bus lane",
  dom_desordem1_escala = "Absence of infrastructure and vegetation",
  dom_desordem2_escala = "Absence of building and green spaces")

outcome_labels <- c(
  aas_inv2 = "Self-rated health (poor vs. good)")

exposure_vars <- names(exposure_labels)
outcome_vars  <- names(outcome_labels)

run_model_row_ct_crude <- function(outcome_var, exp_var) {

  model_formula <- as.formula(
    paste(outcome_var, "~", exp_var))

  model_fit_ct_crude <- glm(
    formula = model_formula,
    data = analysis_data_ct_crude,
    family = poisson(link = "log"),
    weights = normalized_weight)

  robust_vcov <- sandwich::vcovCL(
    model_fit_ct_crude,
    cluster = analysis_data_ct_crude$census_tract)

  lbl_list <- list()
  lbl_list[[exp_var]] <- exposure_labels[[exp_var]]

  tbl_regression(
    model_fit_ct_crude,
    include = all_of(exp_var),
    exponentiate = TRUE,
    label = lbl_list,
    vcov = robust_vcov,
    tidy_fun = broom.helpers::tidy_parameters,
    estimate_fun = function(x) style_number(x, digits = 2),
    pvalue_fun = function(x) style_pvalue(x, digits = 3),
    show_single_row = all_of(exp_var)
  ) %>%
    modify_table_body(
      ~ .x %>%
        mutate(ci = gsub(",", " -", ci)) %>%
        select(-any_of(c("group", "conf.low", "conf.high")))
    ) %>%
    modify_header(
      label = "**Neighborhood indicators**",
      estimate = "**PR**",
      ci = "**95% CI**"
    )
}

table_list_ct_crude <- map(outcome_vars, function(out_v) {
  sub_tables <- map(exposure_vars, ~run_model_row_ct_crude(out_v, .x))
  tbl_stack(sub_tables)})

final_table_ct_crude <- tbl_stack(
  tbls = table_list_ct_crude,
  group_header = outcome_labels[outcome_vars])

gt_table_crude <- final_table_ct_crude %>%
  as_gt()

gt_table_crude$`_source_notes` <- list()

gt_table_crude %>%
  gt::cols_hide(
    columns = matches("^group")
  ) %>%
  gt::cols_width(
    label ~ gt::px(500)
  ) %>%
  gt::tab_options(
    table.width = gt::pct(100)
  ) %>%
  gt::tab_source_note(
    source_note = "Abbreviations: PR = Prevalence Ratio; CI = Confidence Interval. Each association was estimated in a separate Poisson regression model with robust variance clustered at the census tract level.")
```

```{r}
#| label: Minimum adjustment set
#| code-summary: Minimum adjustment set
#| warning: false
#| message: false

analysis_data_ct_minimo <- df %>%
  filter(zona == 1, !is.na(identificador), !is.na(aas_inv2)) %>%
  mutate(
    census_tract = factor(setor),
    normalized_weight = peso_calibrado / mean(peso_calibrado, na.rm = TRUE)
  )

adjustments_minimo <- c(
  "sexo", "faixaetaria", "cor_raca2", "escolaridade2", "rendadompc2corr")

run_model_row_ct_minimo <- function(outcome_var, exp_var) {

  model_formula <- as.formula(
    paste(
      outcome_var, "~", exp_var, "+",
      paste(adjustments_minimo, collapse = " + ")
    )
  )

  model_fit_ct_minimo <- glm(
    formula = model_formula,
    data = analysis_data_ct_minimo,
    family = poisson(link = "log"),
    weights = normalized_weight
  )

  robust_vcov_minimo <- sandwich::vcovCL(
    model_fit_ct_minimo,
    cluster = analysis_data_ct_minimo$census_tract
  )

  lbl_list <- list()
  lbl_list[[exp_var]] <- exposure_labels[[exp_var]]

  table_out <- tbl_regression(
    model_fit_ct_minimo,
    include = all_of(exp_var),
    exponentiate = TRUE,
    label = lbl_list,
    vcov = robust_vcov_minimo,
    tidy_fun = broom.helpers::tidy_parameters,
    estimate_fun = function(x) style_number(x, digits = 2),
    pvalue_fun = function(x) style_pvalue(x, digits = 3),
    show_single_row = all_of(exp_var)
  ) %>%
    modify_table_body(
      ~ .x %>%
        mutate(ci = gsub(",", " -", ci)) %>%
        select(-any_of(c("group", "conf.low", "conf.high")))
    ) %>%
    modify_header(
      label = "**Neighborhood indicators**",
      estimate = "**PR**",
      ci = "**95% CI**")
    return(list(
    model = model_fit_ct_minimo,
    table = table_out
))
}

results_list_ct_minimo <- map(outcome_vars, function(out_v) {
  map(exposure_vars, ~run_model_row_ct_minimo(out_v, .x))
})

models_ct_minimo <- results_list_ct_minimo %>%
  purrr::flatten() %>%
  purrr::map("model")

table_list_ct_minimo <- map(results_list_ct_minimo, function(res_outcome) {
  sub_tables <- map(res_outcome, "table")
  tbl_stack(sub_tables)
})

final_table_ct_minimo <- tbl_stack(
  tbls = table_list_ct_minimo,
  group_header = outcome_labels[outcome_vars]
)

gt_table_minimo <- final_table_ct_minimo %>%
  as_gt()

gt_table_minimo$`_source_notes` <- list()

gt_table_minimo %>%
  gt::cols_hide(
    columns = matches("^group")
  ) %>%
  gt::cols_width(
    label ~ gt::px(500)
  ) %>%
  gt::tab_options(
    table.width = gt::pct(100)
  ) %>%
  gt::tab_source_note(
    source_note = "Abbreviations: PR = Prevalence Ratio; CI = Confidence Interval. Each association was estimated in a separate Poisson regression model with robust variance clustered at the census tract level, adjusted for sex, age group, race/color, education, and household income per capita. Sampling weights were incorporated as normalized weights."
  )
```

```{r}
#| label: VIF
#| code-summary: VIF
#| warning: false
#| message: false

vif_model_minimo <- glm(
  aas_inv2 ~ dom_urbana1_escala + dom_urbana2_escala + dom_transito1_escala + dom_transito2_escala + dom_desordem1_escala + dom_desordem2_escala + sexo + faixaetaria + cor_raca2 + escolaridade2 + rendadompc2corr,
  data = analysis_data_ct_minimo,
  family = poisson(link = "log"))

car::vif(vif_model_minimo)
```

```{r}
#| label: Post-model fit diagnostics
#| code-summary: Post-model fit diagnostics

# Dispersion: ~1 (adequate), >2 (overdispersion), <1 (underdispersion)
# Pearson residuals: < −3 or > 3 (possible outliers)
# Cook’s distance: >1 (influential observations)
# Leverage: higher values indicate greater influence on fitted values


check_model_fit <- function(model) {
  # Pearson dispersion
  dispersion <- sum(residuals(model, type = "pearson")^2) / df.residual(model)
  # Residuals
  res <- residuals(model, type = "pearson")
  # Influence
  cooks <- cooks.distance(model)
  # Leverage
  hat <- hatvalues(model)
  data.frame(
    n = nobs(model),
    dispersion = dispersion,
    min_residual = min(res),
    max_residual = max(res),
    max_cooks = max(cooks, na.rm = TRUE),
    max_leverage = max(hat, na.rm = TRUE)
  )
}

models_ct_minimo <- results_list_ct_minimo %>%
  purrr::flatten() %>%
  purrr::map("model")

diagnostics_ct_minimo <- purrr::map_df(models_ct_minimo, check_model_fit)

diagnostics_ct_minimo
```

```{r}
#| label: Interaction with age
#| code-summary: Interaction with age
#| warning: false
#| message: false

analysis_data_ct_int_age <- df %>%
  filter(zona == 1, !is.na(identificador), !is.na(aas_inv2)) %>%
  mutate(
    census_tract = factor(setor),
    normalized_weight = peso_calibrado / mean(peso_calibrado, na.rm = TRUE)
  )

# Minimum adjustment set
adjustments_int_age <- c(
  "sexo", "cor_raca2", "escolaridade2", "rendadompc2corr")

run_model_row_ct_int_age <- function(outcome_var, exp_var) {

  model_formula <- as.formula(
    paste(
      outcome_var, "~",
      paste0(exp_var, " * faixaetaria"), "+",
      paste(adjustments_int_age, collapse = " + ")
    )
  )

  model_fit_ct_int_age <- glm(
    formula = model_formula,
    data = analysis_data_ct_int_age,
    family = poisson(link = "log"),
    weights = normalized_weight
  )

  robust_vcov_int_age <- sandwich::vcovCL(
    model_fit_ct_int_age,
    cluster = analysis_data_ct_int_age$census_tract
  )

  model_summary <- lmtest::coeftest(
    model_fit_ct_int_age,
    vcov = robust_vcov_int_age
  )

  return(list(
    model = model_fit_ct_int_age,
    summary = model_summary
  ))
}

results_list_ct_int_age <- map(outcome_vars, function(out_v) {
  map(exposure_vars, ~run_model_row_ct_int_age(out_v, .x))
})

models_ct_int_age <- results_list_ct_int_age %>%
  purrr::flatten() %>%
  purrr::map("model")

summaries_ct_int_age <- results_list_ct_int_age %>%
  purrr::flatten() %>%
  purrr::map("summary")

summaries_ct_int_age
```
